---
date: "2017-06-14"
author: "Nima Hejazi"
title: "renjin: speeding up R with Java"
type: "post"
categories: [ "R", "blog", "tools", "productivity", "computing" ]

---

```{r setup, include=FALSE}
library(nima)
```

R is a great, highly flexible language for statistical computing, but it does
suffer greatly from performance issues. As I've steadily increase my use of R, I
quickly became aware that I would have to one day learn to integrate R with a
programming language with better performance, the main choice here being C++. To
integrate R with C++, the `Rcpp` framework (and R package) was created, allowing
for parts of the R code of a given package or project to be re-written in C++
and easily integrated with R. Using `Rcpp` comes with great advantages in terms
of R code performance; however, it obviously requires that one learn C++. I was
about to devote a great deal of time to doing so, when -- fortuitously -- I came
across the rather new [`renjin` project](http://www.renjin.org/). Renjin is a
new (in-development) interpreter for GNU R that relies on the Java Virtual
Machine (JVM) to enhance R's performance. The idea seems to be that it can
eventually serve as a drop-in replacement for GNU R. It also seems that the
`renjin` R package can be used with the standard GNU R interpreter to wrap
existing R code and provide huge performance increases by using the JVM.

## Minimal example

I just thought I would try the example from the `renjin` R package documentation
out, before moving on to more involved examples. Here we go:

```{r, eval=FALSE}
install.packages("https://nexus.bedatadriven.com/content/groups/public/org/renjin/renjin-gnur-package/0.8.2424/renjin-gnur-package-0.8.2424.tar.gz")
library(renjin)
```

```{r}
bigsum <- function(n) {
  sum <- 0
  for(i in seq(from = 1, to = n)) {
    sum <- sum + i
  }
  sum
}

bigsumc <- compiler::cmpfun(bigsum) # GNU R's byte code compiler

system.time(bigsum(1e5))
system.time(bigsumc(1e5))
```

From the above, we notice that directly using the bytecode compiler improves
performance of our `bigsum` function. Maybe `renjin` can help out even more?

```{r, eval=FALSE}
tryCatch(system.time(renjin(bigsum(1e5))), function(e) print("error"))

renjin(bigsum(1e5))
```

Well, it looks like `renjin` isn't quite ready for prime time. It seems to [fail
when acting on bytecode compiled functions (the standard in GNU R
3.4+)](https://twitter.com/dselivanov_/status/874981607635124225).
