---
date: "2017-05-29"
author: "Nima Hejazi"
title: "taking blogdown for a test drive"
type: "post"
categories: [ "R", "blogdown", "blog" ]

---

Recently, I came across the [blogdown R
package](https://bookdown.org/yihui/blogdown/), a variant of RStudio's popular
[bookdown R package](https://bookdown.org/yihui/bookdown/), made by Yihui Xie
and Amber Thomas. Blogdown allows the user to write blog posts with code chunks,
in any of the large variety of languages supported by RMarkdown, allowing for
computationally reproducible writing and programming. It also plays well with
the new static site engine [Hugo](https://gohugo.io/). Here, I'm mostly just
going to take `blogdown` for a spin.

First, let's generate some data and try doing some very simple summary
statistics:


```{r data}
n <- 100
tx_mean <- 25
W <- replicate(3, rnorm(n))
A <- rnorm(n, mean = tx_mean, sd = 2*sqrt(tx_mean))
Y <- as.numeric(A > tx_mean & W[, 1] > 0)
O <- as.data.frame(cbind(Y, A, W))
skimr::skim(O)
```

Look at that! In the above, we generate background covariates (W) based on the
standard Normal distribution, a treatment (A) based on a Normal distribution
with specified mean and standard deviation, and an outcome (Y) based on a
specified structural equation of the form:

$$Y = I(A > 25) \cdot I(W_1 > 0),$$ for n = 100.

Having just recently returned from [ACIC '17](http://causal.unc.edu/acic2017/),
I have causal inference on my mind. You'll notice that in specifying the data
generating processes above, I provided a structural equation for Y -- now, let's
play with marginal structural models (MSMs) just a little bit:

```{r msm}
msm <- glm(Y ~ ., family = "binomial", data = O)
summary(msm)
```

We then project the observed data (O) onto a marginal structural model (MSM) for
the outcome, using a logistic regression with terms for all of the baseline
covariates and the treatment. The `summary` of the resultant `glm` object is
displayed above. Notice that the estimated coefficients for the terms
corresponding to the treatment (A) and first baseline covariate (W_1) are
statistically significant -- recall that these were the variables we used in
specifying the structural equation for Y above.

I think I'll wrap up this experiment by trying some simple plotting:

```{r plots, fig.align='center', fig.cap="Logistic regression of A on Y"}
library(ggplot2)
p <- ggplot(O, aes(A, Y)) + geom_point()
p <- p + stat_smooth(method = "glm", method.args = list(family = "binomial"),
                     formula = y ~ x)
p <- p + theme_minimal() + xlab("A (treatment)") + ylab("Y (outcome)")
print(p)
```

We can plot the relationship between the treatment and outcome using a logistic
regression fit. Pretty awesome that `blogdown` supports such nice graphics so
easily. In my old blog, I had to come up with a custom R script to use `knitr`
and `RMarkdown` to process Rmd files to standard markdown, allowing them to be
rendered on my old [Jekyll](https://jekyllrb.com/) blog. Great to not have to do
that ever again...
