---
date: "2017-12-26"
author: "Nima Hejazi"
title: "sl3: Machine Learning Pipelines for R"
type: "post"
categories: [ "R", "data science", "machine learning", "computing" ]
comments: false
published: false

---

IN PROGRESS:
* https://www.kdnuggets.com/2017/12/managing-machine-learning-workflows-scikit-learn-pipelines-part-1.html
* https://github.com/jeremyrcoyle/sl3/issues/104

Common in the language of modern data science are words such as "munging,"
"massaging," "mining" -- all words denoting the interactive process by which the
analyst extracts some form of deliverable inference from a given data set. These
terms express, among other things, the (often) convoluted process by which a set
of pre-processing and estimation procedures are applied to an input data set in
order to transform said data set into a
["tidy"](http://vita.had.co.nz/papers/tidy-data.html) output data set from which
informative visualizations and summaries may be easily extracted. A formalism
that captures this involved process is that of machine learning _pipelines_. A
_pipeline_, popularized by the [method of the same
name](http://scikit-learn.org/stable/modules/generated/sklearn.pipeline.Pipeline.html)
in Python's [scikit-learn library](http://scikit-learn.org/stable/index.html),
may be thought of as a simple bundle that documents procedures to be applied to
as input data set in a particular order, ultimately resulting in a tidy output
data set.

Recently, the pipeline idiom has made its way into the R programming language,
via the new [`sl3` R package](https://github.com/jeremyrcoyle/sl3). A concrete
understanding of the utility of pipelines is best developed by example -- so,
that's precisely what we'll do! In the following, we'll apply the concept of a
machine learning pipeline to the canonical [iris data set](), combining a series
of learners (machine learning algorithms for estimation/classification) with
Principal Components analysis, a simple pre-processing step.

```{r setup_pkgs, message=FALSE}
library(datasets)
library(tidyverse)
library(data.table)
library(sl3)
```

...

```{r prepare_data, message=FALSE}
data(iris)
iris <- iris %>%
  as_tibble(.)
iris
```

...

```{r make_iris_task}
iris_task <- sl3_Task$new(
  data = iris,
  covariates = colnames(iris)[-5],
  outcome = colnames(iris)[5],
  outcome_type = "categorical"
)
iris_task
```

...

```{r}
```

