---
author: "Nima Hejazi"
categories: [ "R", "data science", "tools", "productivity" ]
tags: [ "R", "data science", "tools", "productivity" ]
date: "2017-05-29"
description: "Trying out RStudio's new blogging framework"
featured: ""
featuredalt: ""
featuredpath: ""
linktitle: ""
title: "Taking blogdown for a test drive"
type: "post"
#comments: true
disableComments: false
published: true
output:
  blogdown::html_page:
    toc: false

---

```{r setup, include=FALSE}
library(ggplot2)
library(skimr)
library(nima)
```

Recently, I came across the [blogdown R
package](https://bookdown.org/yihui/blogdown/), a variant of RStudio's popular
[bookdown R package](https://bookdown.org/yihui/bookdown/), made by Yihui Xie
and Amber Thomas. Blogdown allows the user to write blog posts with code chunks,
in any of the large variety of languages supported by RMarkdown, allowing for
computationally reproducible writing and programming. It also plays well with
the new static site engine [Hugo](https://gohugo.io/). Here, I'm mostly just
going to take `blogdown` for a spin.

First, let's generate some data and try doing some very simple summary
statistics:

```{r data_simulation}
# we're going to be simulating...set seed and constants
set.seed(6142)
n <- 1000
tx_mean <- 25

# generate variables randomly and by structural equation
W <- replicate(3, rnorm(n))
A <- rnorm(n, mean = tx_mean, sd = 2*sqrt(tx_mean))
Y <- as.numeric(A > tx_mean & W[, 1] > 0)
O <- as.data.frame(cbind(Y, A, W))
colnames(O) <- c("Y", "A", paste0("W", seq_len(3)))
head(O)
skim(O)
```

Look at that! In the above, we generate background covariates ($W$) based on the
standard Normal distribution, a treatment ($A$) based on a Normal distribution
with specified mean ($\mu = 25$) and standard deviation
($\sigma = 2 \cdot \sqrt{\mu} = 10$), and an outcome ($Y$) based on a specified
structural equation of the form:

$$Y = I(A > 25) \cdot I(W_1 > 0),$$ for $n = 100$.

Having just recently returned from [ACIC '17](http://causal.unc.edu/acic2017/),
I have causal inference on my mind. You'll notice that in specifying the data
generating processes above, I provided a structural equation for $Y$ -- now,
let's play with marginal structural models (MSMs) just a little bit:

```{r msm}
msm <- glm(Y ~ ., family = "binomial", data = O)
summary(msm)
```

Above, we projected the observed data ($O = (W, A, Y)$) onto a working MSM for
the outcome, using a logistic regression with terms for all of the baseline
covariates and the treatment. From the `summary` of the resultant `glm` object,
we notice that the estimated coefficients for the terms corresponding to the
treatment ($A$) and first baseline covariate ($W_1$) are statistically
significant -- recall that these were the variables we used in specifying the
structural equation for $Y$ above.

I think I'll wrap up this experiment by trying some simple plotting:

```{r plots, fig.align='center', fig.cap="Logistic regression of A on Y"}
p <- ggplot(O, aes(A, Y)) + geom_point()
p <- p + stat_smooth(method = "glm", method.args = list(family = "binomial"),
                     formula = y ~ x)
p <- p + xlab("A (treatment)") + ylab("Y (outcome)") + theme_nima()
print(p)
```

We can plot the relationship between the treatment and outcome using a logistic
regression fit. Pretty awesome that `blogdown` supports such nice graphics so
easily. In my old blog, I had to come up with a custom R script to use `knitr`
to process [R Markdown](http://rmarkdown.rstudio.com/) to standard markdown,
allowing them to be rendered on my old [Jekyll](https://jekyllrb.com/) blog.

